Pure Spin Current Injection in Hydrogenated Graphene Structures

### 2017-08-02 mu/article/article.tex
Deleting coments.

%\documentclass[prb,11pt,tightenlines,twocolumn,aps]{revtex4-1}
\documentclass[floatfix,prb,aps,superscriptaddress,showpacs,11pt,preprint,letterpaper]{revtex4}
\usepackage{amsmath}    % need for subequations 
\usepackage{graphicx}   % need for figures 
\usepackage{verbatim}   % useful for program listings 
\usepackage{color}      % use if color is used in text 
\usepackage{subfigure}  % use for side-by-side figures 
\usepackage{hyperref}   % use for hypertext links, including those to external 
                        % documents and URLs 
\usepackage[dvipsnames]{xcolor}
\usepackage{multirow}
%\usepackage[inline]{showlabels}
% \input{/Users/bms/util/definitions}
\input{/Users/reinaldo/Documents/tex_files/bmsdefinitions.tex}
\def\tama{10cm}
\raggedbottom           % don't add extra vertical space

@@

\begin{abstract}
We present a theoretical study of spin-velocity injection of a pure spin current
induced by  a linearly polarized beam of light that impinges normally on the
surface of  two 50\% hydrogenated noncentrosymmetric 2D graphene structures,
labeled Up and Alt. The hydrogenation opens an energy gap in both structures. We
analyze two possibilities: one in which by fixing the spin along a chosen
direction, the resulting spin-velocity is calculated, and another where we
choose the spin-velocity direction on the surface plane, and calculate its speed
and spin orientation. We do so by changing the energy $\hbar\go$ and
polarization angle $\ga$  of the incoming  beam. The results are calculated
within a full electronic band structure scheme within the DFT in the LDA
approximation. The maxima of the spin-velocity speed are obtained when
$\hbar\go=0.084$\,eV and $\ga=35^\circ$ for Up, and $\hbar\go=0.720$\,eV and
$\ga=150^\circ$ for Alt. We find a speed of 668\,Km/s and 645\,Km/s for Up and
Alt, respectively, when the spin points perpendicularly to the surface. Also,
the response is maximized by fixing the spin-velocity direction along a high
symmetry Cartesian axis, obtaining a speed of 688\,Km/s with the spin  pointing
at 13$^\circ$ from the surface normal, for Up, and 906\,Km/s and the spin
pointing at 60$^\circ$ from the surface normal, for Alt. These speed values are
larger than those of bulk semiconductors, like CdSe and GaAs, thus making the
hydrogenated graphene structures excellent candidates for spintronics
applications.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%               %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  T H E O R Y  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%               %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%% Theory: Spin velocity %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%BMSD

@@

\hat K^{\rma\rmb}=\frac{1}{2}\left( \hat v^\rma \hat S^\rmb 
+\hat  S^\rmb \hat v^\rma\right) 

@@

where $\hat\bfv=[\hat\bfr,\hat H_0]/i\hbar$ is the velocity operator, with
$\hat \bfr$ being the position operator and $\hat H_0$ the unperturbed
ground state Hamiltonian;
the Roman superscripts  indicate Cartesian coordinates. 
To obtain the expectation value of 
$\hat K^{\rma\rmb}$, we use the length gauge for the perturbing
Hamiltonian, written as

@@

\hat H_{\text{p}}=-e\hat\bfr\cdot\bfE(t)

@@

\bfE(t)= \bfE(\go)e^{-i\go t} + \bfE^*(\go)e^{i\go t}

@@

In order to 
calculate the response of the system to $\bfE(t)$, one needs to
take into account the excited coherent superposition
of the spin-split conduction bands inherent to the 
noncentrosymmetric 
semiconductors considered in this work.
%Although, this splitting is small [15,16], is required to obtain the
%correct results\cite{nastosPRB07}.
% The conduction bands in the 
% noncentrosymmetric  
% semiconductors  
% are spin split by a small amount,[15,16] typically
%   smaller than the energy width of the laser pulse, and so the pulse
%   excites a coherent superposition of the two conduction bands. Even
%   for very long pulses with narrow energy widths, dephasing effects
%   lead to an energy width of the bands large enough that spin-split
%   states can become quasidegenerate. 
To include the coherences, we follow Ref.~\onlinecite{nastosPRB05} and
use a multiple
scale approach that solves the equation of motion for the single
particle density matrix $\gr_{mn}(\bfk;t)$, leading to

@@

&\frac{\partial \rho_{cc'}(\bfk)}{\partial t} =
\frac{e^{2}E^{\mathrm{a}}(\omega)E^{\mathrm{b*}}(\omega)}
{i \hbar^{2}}
\sum_{v}r^{\mathrm{a}}_{cv}(\bfk) r^{\mathrm{b}}_{vc'}(\bfk)
\left( \frac{1}{\omega - \omega_{c'v}(\bfk) - i \epsilon} 
- 
\frac{1}{\omega - \omega_{cv}(\bfk) + i \epsilon} \right)

@@

where we assumed that the conduction bands $c$ and $c'$
are quasidegenerate  states, and we take $\ge\to 0$ at the end of the
calculation.  The spin-splitting of the valence ($v$) bands is very
small, and is neglected throughout this work.\cite{nastosPRB07}
% 􏱈 are close to  
% one another, and that the pulse is short enough so that the energy
% width overlaps the two bands, the equation of motion can be
% solved. Leaving the details of the derivation to Ap- 
% pendix A, the result for the off-diagonal component 􏰠 􏱈, c  
%%%BMSU 
The matrix elements of any operator $\calo$ are given by
$\calo_{nm}(\bfk)=\bra{n\bfk}\hat\calo\ket{m\bfk}$,
where 
$H_{0}|n\mathbf{k}\rangle = \hbar \omega_{n}(\mathbf{k})|n\mathbf{k}\rangle$ 
with $\hbar \omega_{n}(\mathbf{k})$ being the energy of the
electronic band $n$ at point $\mathbf{k}$ in the irreducible Brillouin zone
(IBZ),  $|n\mathbf{k}\rangle$ is the Bloch state, and 
$\go_{nm}(\bfk)=\go_{n}(\bfk)-\go_{m}(\bfk)$.
Using 
$\mathcal{O} = \mathrm{Tr}(\hat{\rho}\hat{\mathcal{O}})$
for the expectation value of an observable $\mathcal{O}$, 
where $\mathrm{Tr}$ denotes the trace, we obtain

@@

\mathcal{O} 
%=&
%\int \frac{d^{3}k}{8\pi^{3}}\sum_{c} \langle c \mathbf{k} 
%| \hat{\rho} \hat{\mathcal{O}} |
%c\mathbf{k} \rangle \nonumber \\
=& 
\int \frac{d^{3}k}{8\pi^{3}} \sum_{cc'} \rho_{cc'}(\mathbf{k}) 
\mathcal{O}_{c'c}(\mathbf{k}),

@@

where we used 
the closure
relationship $\sum_{n}|n\mathbf{k}\rangle \langle n\mathbf{k}| = 1$,
where $n=v,c$,
and the fact that $\rho_{vn}(\bfk)=\rho_{nv}(\bfk)=0$ for $n=c,c'$. 
Therefore, 
using  Eqs. \eqref{z.4} and \eqref{z.5},
the
rate of change of $\mathcal{O}$,
$\dot{\mathcal{O}} 
\mathrm{Tr} \left( \dot{\hat\rho} \hat\calo \right)
$,
 is given by

@@

\dot{\mathcal{O}} 
&=\frac{e^{2}}{i\hbar^{2}} \int \frac{d^{3}k}{8\pi^{3}} 
\sum'_{cc'} \mathcal{O}_{c'c}(\bfk) 
r^{\mathrm{a}}_{cv}(\bfk)  r^{\mathrm{b}}_{vc'}(\bfk)  
\left( \frac{1}{\omega - \omega_{c'v}(\bfk)  - i\epsilon} - 
\frac{1}{\omega - \omega_{cv}(\bfk)  + i\epsilon} \right)

@@

Replacing  $\hat\calo \rightarrow \hat{K}^{\mathrm{ab}}$, in 
the above expression, one can show that

@@

\dot{K}^{\mathrm{ab}}(\omega) =
\mu^{\mathrm{abcd}}(\omega)

@@

  \delta(\omega-\omega_{cv}(\bfk) 
\mathrm{Re} & \left[ K^{\mathrm{ab}}_{cc'}(\bfk) 

@@

r^{\mathrm{c}}_{vc'}(\bfk)   
r^{\mathrm{d}}_{cv }(\bfk)  +

@@

$
K^{\mathrm{ab}}_{nm}(\mathbf{-k}) = K^{\mathrm{ab*}}_{nm}(\mathbf{k}) 
$
which follows from time-reversal invariance. The prime symbol $'$ in the sum means that
$c$ and $c'$ are quasi degenerate states, and the sum only covers these states.
Since $\mu^{\mathrm{abcd}}(\omega)$ is real, we have that
$\mu^{\mathrm{abcd}}(\omega) =
\mu^{\mathrm{abdc}}(\omega)$. 
We remark that Eq.~\eqref{eq:mu} 
is the same as Eq. (3) of Bhat et al.\cite{bhatPRL05},
obtained by using the semiconductor optical Bloch equations.
Using the closure relation,

@@

K^{\mathrm{ab}}_{cc'}(\mathbf{k}) = \frac{1}{2}
\sum_{l=v,c}
\left(v^{\mathrm{a}}_{cl}(\mathbf{k})S^{\mathrm{b}}_{lc'}(\mathbf{k})
+S^{\mathrm{b}}_{cl}(\mathbf{k}) v^{\mathrm{a}}_{lc'}(\mathbf{k})

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\subsection{Spin velocity injection} % (fold)
%\label{sec:theory-pure_spin_current}
Now, we define the spin velocity injection (SVI) as
\begin{equation}\label{eq:vab-w}
\mathcal{V}^{\mathrm{ab}}(\omega) \equiv
\frac{\dot{K}^{\mathrm{ab}}(\omega)}{(\hbar/2) \dot{n}(\omega)},
\end{equation}  
which gives the velocity, along direction $\mathrm{a}$, at which the spin
moves in a polarized state along direction $\mathrm{b}$. The carrier injection rate
$\dot n(\go)$ is written as,\cite{nastosPRB05}
\begin{equation}
\dot{n}(\omega) =
\xi^{\mathrm{ab}}(\omega) E^{c }(\omega) E^{d*}(\omega)
\label{eq:dotn}
\end{equation}
where the tensor 
\begin{equation}\label{eq:xi}
\begin{aligned}
\xi^{\mathrm{ab}}(\omega)
&
=
\frac{2\pi e^{2}}{\hbar^{2}} \int 
\frac{d^{3}k}{8 \pi^{3}}
 \sum_{vc}
r^{\mathrm{a}}_{vc'}(\bfk)  
r^{\mathrm{b}}_{cv }(\bfk)  
\delta(\omega-\omega_{cv}(\bfk) 
, 
\end{aligned}
\end{equation}
is related to the imaginary part of the linear optical 
response tensor by
$\mathrm{Im}[\ge^{\rma\rmb}(\go)]=2\pi\ge_0\hbar\xi^{\rma\rmb}(\go)$.

The function $\calv^{\rma\rmb}(\go)$ allows us to quantify two very
important aspects of PSC. On one hand, we can fix the spin direction
along $\mathrm{b}$  
and calculate the resulting electron velocity.  
On the other hand, we can fix the velocity of the electron
along $\mathrm{b}$ and study the resulting direction along which the
spin is polarized.
To this end, the added advantage of  2D structures, besides choosing
them as noncentrosymmetric, is that we can use an incoming linearly
polarized beam of light
at normal incidence, and use the  direction of the polarized  electric
field to control $\calv^{\rma\rmb}(\go)$.
Indeed, writing 
$\bfE(\omega) = E_0(\omega)(\cos\ga\,\hat\bfx+\sin\ga\,\hat\bfy)$,
where $\alpha$ is the polarization angle, we obtain from
 Eq. \eqref{eq:vab-w}
that
\begin{widetext}
\begin{align}
\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)
&= 
\frac{2}{\hbar\xi(\go)}
\left(\mu^{\mathrm{abxx}}(\omega)\cos^{2}\alpha + 
\mu^{\mathrm{abyy}}(\omega)\sin^{2}\alpha + 
\mu^{\mathrm{abxy}}(\omega)\sin 2\alpha\right)
,
\label{eq:vab-aw}
\end{align}
\end{widetext}
for the structures chosen in this article,
$\xi^{\mathrm{xx}}(\go)=\xi^{\mathrm{yy}}(\go)\equiv\xi(\go)$, and $\xi^{\mathrm{xy}}(\go)=0$.
Now, we formalize our two options for $\calv^{\rma\rmb}(\go)$.

@@

Analyzing the SVI, Eq. \eqref{eq:vab-aw}, we define the magnitude of
the electron velocity on the plane of the structure,
with the spin polarized along the $\mathrm{b}$ direction as
\begin{equation}
\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)
\equiv
\sqrt{
[\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)]^{2}\ +
[\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)]^{2}\ 
}, 
\label{eq:vs-mag}
\end{equation}
and define the angle at which the velocity is directed on the $xy$ plane as
\begin{equation}
\gamma_{\gs^\mathrm{b}} (\omega,\alpha)
=
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)} \right)
.
\label{eq:gamma-ang}
\end{equation}
% where this angle is measured in the counter-clockwise direction from the
% positive $x$. 
We also define two special angles
\begin{equation}
\gamma_{\gs^\mathrm{b}}^\parallel(\omega,\alpha) = \alpha, 
\label{eq:gamma-par} 
\end{equation}
and
\begin{equation}
\gamma_{\gs^\mathrm{b}}^\perp(\omega,\alpha) = \alpha \pm 90^{\circ},
\label{eq:gamma-perp}
\end{equation}
corresponding to the electron velocity being parallel or perpendicular
to the incoming polarization, 
respectively. The subscript $\gs^\rmb$ denotes the spin along $\rmb$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%% Theory: Fixing vel %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Fixing the calculated velocity along $\rma=x$ or $\rma=y$,
we define its corresponding magnitude as
\begin{align}
\mathcal{V}_{\mathrm{a}}(\omega,\alpha) \equiv 
\sqrt { 
[\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{az}}(\omega,\alpha)]^{2} 
},
\label{eq:vv-mag}
\end{align}
from where we see that the spin would be oriented
in the $xyz$ system's coordinates
according to a polar angle
\begin{align}
\theta_{\mathrm{a}}  (\omega,\alpha)
= 
\cos^{-1} \left( \frac{\mathcal{V}^{\mathrm{az}}(\omega,\alpha)}
{\mathcal{V}_{\mathrm{a}}(\omega,\alpha)} \right),
 \quad 0 \leq \theta \leq \pi, 
\label{eq:polar-ang}
\end{align}
and an azimuthal angle
\begin{align}
\varphi_{\mathrm{a}} (\omega,\alpha)
=
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)} \right),
\quad 0 \leq \varphi \leq 2\pi.
\label{eq:azimuthal-ang} 
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%   R E S U L T S   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%% Res: mu & V^{ab} comparison %%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\centering
\includegraphics[width=\tama]{plots/2-vab-str-comp}
\caption{$\mathcal{V}^{\mathrm{ab}}(\go,\ga)$ 
  vs. $\hbar\go$, for the values of $\ga$ that maximize the signal. The low
  energy regions for the Alt and Up structures start at the energy gap of each
  structure, while in the high energy regions, the values of
  $\mathcal{V}^{\mathrm{ab}}(\go,\ga)$ are also very large. The bottom panel
  corresponds to CdSe and GaAs. }
\label{fig:vab-str-comp}
\end{figure}

>>>>

%\documentclass[prb,11pt,tightenlines,twocolumn,aps]{revtex4-1}
\documentclass[floatfix,prb,aps,superscriptaddress,showpacs,11pt,preprint,letterpaper]{revtex4}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{color}
\usepackage{subfigure}
\usepackage{hyperref}
\usepackage[dvipsnames]{xcolor}
\usepackage{multirow}
% \input{/Users/bms/util/definitions}
%\input{/Users/reinaldo/Documents/tex_files/bmsdefinitions.tex}
\def\tama{10cm}
\raggedbottom

@@

\begin{abstract}
We present a theoretical study of spin-velocity injection of a pure spin current
induced by  a linearly polarized beam of light that impinges normally on the
surface of  two 50\% hydrogenated noncentrosymmetric 2D graphene structures,
labeled Up and Alt. The hydrogenation opens an energy gap in both structures. We
analyze two possibilities: one in which by fixing the spin along a chosen
direction, the resulting spin-velocity is calculated, and another where we
choose the spin-velocity direction on the surface plane, and calculate its speed
and spin orientation. We do so by changing the energy $\hbar\omega$ and
polarization angle $\alpha$  of the incoming  beam. The results are calculated
within a full electronic band structure scheme within the DFT in the LDA
approximation. The maxima of the spin-velocity speed are obtained when
$\hbar\omega = 0.084$\,eV and $\alpha=35^\circ$ for Up, and $\hbar\omega =
0.720$\,eV and $\alpha=150^\circ$ for Alt. We find a speed of 668\,Km/s and
645\,Km/s for Up and Alt, respectively, when the spin points perpendicularly to
the surface. Also, the response is maximized by fixing the spin-velocity
direction along a high symmetry Cartesian axis, obtaining a speed of 688\,Km/s
with the spin  pointing at 13$^\circ$ from the surface normal, for Up, and
906\,Km/s and the spin pointing at 60$^\circ$ from the surface normal, for Alt.
These speed values are larger than those of bulk semiconductors, like CdSe and
GaAs, thus making the hydrogenated graphene structures excellent candidates for
spintronics applications.
\end{abstract}

@@

@@

@@

@@

\hat K^{\mathrm{a}\mathrm{b}} = 
\frac{1}{2}\left( \hat v^\mathrm{a} \hat S^\mathrm{b} 
+\hat  S^\mathrm{b} \hat v^\mathrm{a}\right) 

@@

where $\hat{\mathbf v}=[\hat{\mathbf r},\hat H_0]/i\hbar$ is the velocity
operator, with $\hat {\mathbf r}$ being the position operator and $\hat H_0$ the
unperturbed ground state Hamiltonian; the Roman superscripts  indicate Cartesian
coordinates. To obtain the expectation value of $\hat K^{\mathrm{a}\mathrm{b}}$,
we use the length gauge for the perturbing Hamiltonian, written as

@@

\hat H_{\text{p}}=-e\hat{\mathbf r}\cdot{\mathbf E}(t)

@@

{\mathbf E}(t) = 
{\mathbf E}(\omega)e^{-i\omega t} + {\mathbf E}^*(\omega)e^{i\omega t}

@@

In order to calculate the response of the system to ${\mathbf E}(t)$, one needs
to take into account the excited coherent superposition of the spin-split
conduction bands inherent to the noncentrosymmetric semiconductors considered in
this work.
%Although, this splitting is small [15,16], is required to obtain the correct
%results\cite{nastosPRB07}. The conduction bands in the noncentrosymmetric
%semiconductors are spin split by a small amount,[15,16] typically smaller than
%the energy width of the laser pulse, and so the pulse excites a coherent
%superposition of the two conduction bands. Even for very long pulses with
%narrow energy widths, dephasing effects lead to an energy width of the bands
%large enough that spin-split states can become quasidegenerate.
To include the coherences, we follow Ref.~\onlinecite{nastosPRB05} and use a
multiple scale approach that solves the equation of motion for the single
particle density matrix ${\rho}_{mn}({\mathbf k};t)$, leading to

@@

&\frac{\partial \rho_{cc'}({\mathbf k})}{\partial t} =
\frac{e^{2}E^{\mathrm{a}}(\omega)E^{\mathrm{b*}}(\omega)}{i \hbar^{2}}
\sum_{v}r^{\mathrm{a}}_{cv}({\mathbf k}) r^{\mathrm{b}}_{vc'}({\mathbf k})
\left( \frac{1}{\omega - \omega_{c'v}({\mathbf k}) - i \epsilon} - 
\frac{1}{\omega - \omega_{cv}({\mathbf k}) + i \epsilon} \right)

@@

where we assumed that the conduction bands $c$ and $c'$ are quasidegenerate
states, and we take $\epsilon\to 0$ at the end of the calculation.  The spin-
splitting of the valence ($v$) bands is very small, and is neglected throughout
this work.\cite{nastosPRB07} The matrix elements of any operator ${\cal O}$ are
given by ${\cal O}_{nm}({\mathbf k})=\langle n{\mathbf k}|\hat{\cal O}|
m{\mathbf k}\rangle$, where $H_{0}|n{\mathbf k}\rangle = \hbar
\omega_{n}({\mathbf k})|n{\mathbf k}\rangle$ with $\hbar \omega_{n}({\mathbf
k})$ being the energy of the electronic band $n$ at point ${\mathbf k}$ in the
irreducible Brillouin zone (IBZ),  $|n{\mathbf k}\rangle$ is the Bloch state,
and $\omega_{nm}({\mathbf k})=\omega_{n}({\mathbf k})-\omega_{m}({\mathbf k})$.
Using $\mathcal{O} = \mathrm{Tr}(\hat{\rho}\hat{\mathcal{O}})$ for the
expectation value of an observable $\mathcal{O}$, where $\mathrm{Tr}$ denotes
the trace, we obtain

@@

\mathcal{O} = & 
\int \frac{d^{3}k}{8\pi^{3}} \sum_{cc'} \rho_{cc'}({\mathbf k}) 
\mathcal{O}_{c'c}({\mathbf k}),

@@

where we used the closure relationship $\sum_{n}|n{\mathbf k}\rangle \langle
n{\mathbf k}| = 1$, where $n=v,c$, and the fact that $\rho_{vn}({\mathbf
k})=\rho_{nv}({\mathbf k})=0$ for $n=c,c'$. Therefore, using  Eqs. \eqref{z.4}
and \eqref{z.5}, the rate of change of $\mathcal{O}$, $\dot{\mathcal{O}} =
\mathrm{Tr} \left( \dot{\hat\rho} \hat{\cal O} \right)$, is given by

@@

\dot{\mathcal{O}} &=
\frac{e^{2}}{i\hbar^{2}} \int \frac{d^{3}k}{8\pi^{3}} 
\sum'_{cc'} \mathcal{O}_{c'c}({\mathbf k}) 
r^{\mathrm{a}}_{cv}({\mathbf k})  r^{\mathrm{b}}_{vc'}({\mathbf k})  
\left( \frac{1}{\omega - \omega_{c'v}({\mathbf k})  - i\epsilon} - 
\frac{1}{\omega - \omega_{cv}({\mathbf k})  + i\epsilon} \right)

@@

Replacing $\hat{\cal O} \rightarrow \hat{K}^{\mathrm{ab}}$, in the above
expression, one can show that

@@

\dot{K}^{\mathrm{ab}}(\omega) = \mu^{\mathrm{abcd}}(\omega)

@@

\delta(\omega-\omega_{cv}({\mathbf k}) 
\mathrm{Re} & \left[ K^{\mathrm{ab}}_{cc'}({\mathbf k}) 

@@

r^{\mathrm{c}}_{vc'}({\mathbf k})   
r^{\mathrm{d}}_{cv }({\mathbf k})  +

@@

$K^{\mathrm{ab}}_{nm}({\mathbf -k}) = K^{\mathrm{ab*}}_{nm}({\mathbf k})$ which
follows from time-reversal invariance. The prime symbol $'$ in the sum means
that $c$ and $c'$ are quasi degenerate states, and the sum only covers these
states. Since $\mu^{\mathrm{abcd}}(\omega)$ is real, we have that
$\mu^{\mathrm{abcd}}(\omega) = \mu^{\mathrm{abdc}}(\omega)$. We remark that
Eq.~\eqref{eq:mu} is the same as Eq. (3) of Bhat et al.\cite{bhatPRL05},
obtained by using the semiconductor optical Bloch equations. Using the closure
relation,

@@

K^{\mathrm{ab}}_{cc'}({\mathbf k}) = 
\frac{1}{2} \sum_{l=v,c}
\left(v^{\mathrm{a}}_{cl}({\mathbf k})S^{\mathrm{b}}_{lc'}({\mathbf k})
+S^{\mathrm{b}}_{cl}({\mathbf k}) v^{\mathrm{a}}_{lc'}({\mathbf k})

@@

Now, we define the spin velocity injection (SVI) as
\begin{equation}\label{eq:vab-w}
\mathcal{V}^{\mathrm{ab}}(\omega) \equiv
\frac{\dot{K}^{\mathrm{ab}}(\omega)}{(\hbar/2) \dot{n}(\omega)},
\end{equation}  
which gives the velocity, along direction $\mathrm{a}$, at which the spin moves
in a polarized state along direction $\mathrm{b}$. The carrier injection rate
$\dot n(\omega)$ is written as,\cite{nastosPRB05}
\begin{equation}
\dot{n}(\omega) =
\xi^{\mathrm{ab}}(\omega) E^{c }(\omega) E^{d*}(\omega)
\label{eq:dotn}
\end{equation}
where the tensor 
\begin{equation}\label{eq:xi}
\xi^{\mathrm{ab}}(\omega) =
\frac{2\pi e^{2}}{\hbar^{2}} \int 
\frac{d^{3}k}{8 \pi^{3}} \sum_{vc}
r^{\mathrm{a}}_{vc'}({\mathbf k})  
r^{\mathrm{b}}_{cv }({\mathbf k})  
\delta(\omega-\omega_{cv}({\mathbf k}) 
, 
\end{equation}
is related to the imaginary part of the linear optical response tensor by
$\mathrm{Im} [\epsilon^{\mathrm{a}\mathrm{b}}(\omega)] =
2\pi\epsilon_0\hbar\xi^{\mathrm{a}\mathrm{b}}(\omega)$.

The function ${\cal V}^{\mathrm{a}\mathrm{b}}(\omega)$ allows us to quantify two
very important aspects of PSC. On one hand, we can fix the spin direction along
$\mathrm{b}$ and calculate the resulting electron velocity. On the other hand,
we can fix the velocity of the electron along $\mathrm{b}$ and study the
resulting direction along which the spin is polarized. To this end, the added
advantage of  2D structures, besides choosing them as noncentrosymmetric, is
that we can use an incoming linearly polarized beam of light at normal
incidence, and use the  direction of the polarized  electric field to control
${\cal V}^{\mathrm{a}\mathrm{b}}(\omega)$. Indeed, writing ${\mathbf E}(\omega)
= E_0(\omega)(\cos\alpha\,\hat{\mathbf x}+\sin\alpha\,\hat{\mathbf y})$, where
$\alpha$ is the polarization angle, we obtain from Eq. \eqref{eq:vab-w} that
\begin{widetext}
\begin{equation}
\mathcal{V}^{\mathrm{ab}}(\omega,\alpha) = 
\frac{2}{\hbar\xi(\omega)}
\left(\mu^{\mathrm{abxx}}(\omega)\cos^{2}\alpha + 
\mu^{\mathrm{abyy}}(\omega)\sin^{2}\alpha + 
\mu^{\mathrm{abxy}}(\omega)\sin 2\alpha\right)
,
\label{eq:vab-aw}
\end{equation}
\end{widetext}
for the structures chosen in this article, $\xi^{\mathrm{xx}}(\omega) =
\xi^{\mathrm{yy}}(\omega)\equiv\xi(\omega)$, and $\xi^{\mathrm{xy}}(\omega)=0$.
Now, we formalize our two options for ${\cal V}^{\mathrm{a}\mathrm{b}}(\omega)$.

@@

Analyzing the SVI, Eq. \eqref{eq:vab-aw}, we define the magnitude of the
electron velocity on the plane of the structure, with the spin polarized along
the $\mathrm{b}$ direction as
\begin{equation}
\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha) \equiv
\sqrt{
[\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)]^{2}\ +
[\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)]^{2}\ 
}, 
\label{eq:vs-mag}
\end{equation}
and define the angle at which the velocity is directed on the $xy$ plane as
\begin{equation}
\gamma_{\sigma^\mathrm{b}} (\omega,\alpha) =
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)} \right)
.
\label{eq:gamma-ang}
\end{equation}
We also define two special angles
\begin{equation}
\gamma_{\sigma^\mathrm{b}}^\parallel(\omega,\alpha) = \alpha, 
\label{eq:gamma-par} 
\end{equation}
and
\begin{equation}
\gamma_{\sigma^\mathrm{b}}^\perp(\omega,\alpha) = \alpha \pm 90^{\circ},
\label{eq:gamma-perp}
\end{equation}
corresponding to the electron velocity being parallel or perpendicular to the
incoming polarization, respectively. The subscript $\sigma^\mathrm{b}$ denotes
the spin along $\mathrm{b}$.

@@

Fixing the calculated velocity along $\mathrm{a}=x$ or $\mathrm{a}=y$,
we define its corresponding magnitude as
\begin{align}
\mathcal{V}_{\mathrm{a}}(\omega,\alpha) \equiv 
\sqrt { 
[\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{az}}(\omega,\alpha)]^{2} 
},
\label{eq:vv-mag}
\end{align}
from where we see that the spin would be oriented in the $xyz$ system's
coordinates according to a polar angle
\begin{equation}
\theta_{\mathrm{a}}  (\omega,\alpha) = 
\cos^{-1} \left( \frac{\mathcal{V}^{\mathrm{az}}(\omega,\alpha)}
{\mathcal{V}_{\mathrm{a}}(\omega,\alpha)} \right),
 \quad 0 \leq \theta \leq \pi, 
\label{eq:polar-ang}
\end{equation}
and an azimuthal angle
\begin{equation}
\varphi_{\mathrm{a}} (\omega,\alpha) =
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)} \right),
\quad 0 \leq \varphi \leq 2\pi.
\label{eq:azimuthal-ang} 
\end{equation}

@@

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{plots/2-vab-str-comp}
\caption{$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ 
  vs. $\hbar\omega$, for the values of $\alpha$ that maximize the signal. The
  low energy regions for the Alt and Up structures start at the energy gap of
  each structure, while in the high energy regions, the values of
  $\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ are also very large. The bottom
  panel corresponds to CdSe and GaAs. }
\label{fig:vab-str-comp}
\end{figure}
